This study focuses on the Internet of Things (IoT) gateway; a common middleware
solution that bridges the gap between physical sensors and devices to internet
applications. There is a shown interest in understanding the characteristics of
different types of gateway architectures both from the research field and the
industry, particularly the IT-consulting firm Attentec in Linköping, Sweden.
Apart from this general approach a study has also been made on the open source
C library \textit{libuv}, used in the common web runtime engine
\textit{NodeJS}. The library has been used to study how asynchronous I/O
operations can be used to improve the IoT gateway performance.

Built on top of task management theory, a set of three general architectural
approaches are identified. Common internal and external properties are
identified based on state-of-the-art gateway implementations found in the
industry and professional guidance. All of these mentioned variables are taken
into account when a general gateway implementation is developed that is
proposed to mimic any architectural level implementation of the gateway. A set
of performance tests are conducted on the implementation to observe how
different configurations of the gateway affect throughput and response time of
data transmitted from simulated devices. The results show that the properties
of the gateway do affect throughput and response time significantly and that
libuv overall helps implement one of the best performing gateway
configurations.

